---
title: "Light / Fluo / Lon - Lat data correction"
author: "Claude Cugerone"
date: "2025-05-22"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, include=FALSE}
rm(list = ls())
gc()             # garbage collector

library(ncdf4) 
library(dplyr)
library(oce)
library(zoo)
library(here)
```
#  GLOBAL VARIABLES
##  Global variables & Paths
```{r variables}
MAX_CHLA_DEPTH <- 175 # define depth at which to consider dark signal (CHLA "absolute zero")
MAX_DEPTH <- 200
SURFACE_NOT_NAN <- 4:10
idprofile <- 1 # Index of the profile displayed on individual plots

ROOT_PROJ <- here() # project root

DATA_FOLDER <- "oceanographic_data"
DATA_FOLDER_CORRECTED <- "oceanographic_data_corrected"
DEPLOY_FOLDER <- "wd11" # Lines to modify to change the SealOfApproval deployment
SEAL_NAME <- paste0(DEPLOY_FOLDER,"-913-18") # ...or the individual SealOfApproval
SEAL_FILE <- paste0(SEAL_NAME, "_hr2_prof.nc") # wd20 in hr1, others in hr2
NC_PATH <- file.path(ROOT_PROJ, DATA_FOLDER, DEPLOY_FOLDER, SEAL_FILE)
RSD_PATH <- file.path(ROOT_PROJ, DATA_FOLDER_CORRECTED, DEPLOY_FOLDER, SEAL_NAME)

BATHY_FOLDER <- "bathy"
FINE_TOPO_FILE <- "gvdem100v3/w001001.adf"
LARGE_TOPO_FILE <- "gebco_2023_n-50.0_s-79.0_w90.0_e180.0.tif"
BATHY_PATH_FINE <- file.path(ROOT_PROJ, BATHY_FOLDER, FINE_TOPO_FILE)
BATHY_PATH_LARGE <- file.path(ROOT_PROJ, BATHY_FOLDER, LARGE_TOPO_FILE)

FAST_ICE_FOLDER <- "fast_ice"
FAST_ICE_FILE <- "mertz_sara_akiko_19.nc" # SealOfApproval
FAST_ICE_PATH <- file.path(ROOT_PROJ, FAST_ICE_FOLDER, FAST_ICE_FILE)

AMSR_FOLDER <- "amsr2"
AMSR_FOLDER_YEAR <- "2019" # SealOfApproval
AMSR_FOLDER_PATH <- file.path(ROOT_PROJ,AMSR_FOLDER,AMSR_FOLDER_YEAR)

```

## NetCDF Extract
```{r nc-code}
nc_data <- nc_open(NC_PATH)
lon   <- ncvar_get(nc_data, "LONGITUDE")
lat   <- ncvar_get(nc_data, "LATITUDE")
raw_chla  <- ncvar_get(nc_data, "CHLA")
raw_light <- ncvar_get(nc_data, "LIGHT")
temperature  <- ncvar_get(nc_data, "TEMP")
sal   <- ncvar_get(nc_data, "PSAL")
pres  <- ncvar_get(nc_data, "PRES")
juld  <- ncvar_get(nc_data, "JULD")
nc_close(nc_data)
```

# Date and coordinates correction
## Date formatting
```{r dataframe}
# Vectorization of depth
pres_vec <- pres[, 1]

# Convert Julian date to UTC
dateTime <- as.POSIXct(juld * 86400, origin = "1950-01-01", tz = "UTC")
dateTime <- round(dateTime, units = "mins")  # round to the nearest minute
```

## Correcting longitude and latitude
```{r corrected lon lat}
# Load diagnostic CSV and convert DATE column to POSIXct
diag <- read.csv("filtered_dives_wd_new.csv", stringsAsFactors = FALSE)
diag$date <- as.POSIXct(diag$DATE, format="%Y-%m-%dT%H:%M")

# Extract the individual of interest and sort by date
df_wd <- diag %>% filter(REF == SEAL_NAME)
df_wd <- df_wd %>% arrange(date)

# Convert dates to numeric for interpolation
t1 <- as.numeric(df_wd$date)
t2 <- as.numeric(as.POSIXct(dateTime))

# Interpolate longitude and latitude at the measurement times
cols <- c("LON", "LAT")
interp_results <- lapply(cols, function(col) {
  approx(x = t1, y = df_wd[[col]], xout = t2, rule = 2)$y
})

names(interp_results) <- paste0(cols, "_corr")
df_interp <- as.data.frame(interp_results)

```

### Plotting original and corrected coordinates
```{r}

# df_original <- data.frame(lon = lon, lat = lat)
# 
# ggplot() +
#   #geom_point(data = df_original, aes(x = lon, y = lat, color = "Original"), alpha = 0.4, size = 2) +
#   geom_point(data = df_interp, aes(x = LON_corr, y = LAT_corr, color = "Interpolated"), alpha = 0.4, size = 2) +
#   scale_color_manual(name = "Type de données", values = c("Original" = "blue", "Interpolated" = "red")) +
#   labs(title = "Corrected Lat/Lon", x = "Longitude", y = "Latitude") +
#   # ylim(-70, -63) +
#   theme_minimal()
```

```{r}
lon <- df_interp$LON_corr
lat <- df_interp$LAT_corr
```

# MLD
```{r MLD}
# Function to find the first valid index
first_valid_index <- function(condition)
{
  idx <- which(condition)
  if (length(idx) > 0) return(idx[1]) else return(NA)
}

n_profiles <- ncol(sal)
n_levels <- nrow(sal)
mld_vec <- rep(NA, n_profiles)
densThreshold <- 0.03
refDepth <- 10
minProfileStartDepth <- 20

# Find the first valid depths (on the temperature profile)
profileStartDepth <- rep(NA, n_profiles)
for (i in 1:n_profiles) {
  firstGood <- first_valid_index(!is.na(temperature[, i]))
  if (!is.na(firstGood) && !is.na(pres[firstGood, i])) {
    profileStartDepth[i] <- pres[firstGood]
  }
}
validProfiles <- which(profileStartDepth <= minProfileStartDepth)

# Loop over valid profiles
for (i in validProfiles) {
  S <- sal[, i]
  T <- temperature[, i]
  P <- pres[, i]

  good <- !is.na(S) & !is.na(T) & !is.na(P)
  if (sum(good) == 0) next  # skip profiles that are completely empty

  # Compute potential density
  ctd <- as.ctd(salinity = S[good], temperature = T[good], pressure = P[good])
  # sigma0 <- swSigma0(ctd@data$salinity, ctd@data$temperature, ctd@data$pressure)
  sigma0 <- swSigma0(ctd@data$salinity, ctd@data$temperature, 0)

  full_sigma0 <- rep(NA, n_levels)
  full_sigma0[which(good)] <- sigma0

  # Find the reference index (first >= 10 m)
  refIdx <- first_valid_index(ctd@data$pressure >= refDepth)

  if (!is.na(refIdx) && !is.na(sigma0[refIdx])) {
    refSigma0 <- sigma0[refIdx]
    # Density difference at all valid levels
    deltaSigma <- full_sigma0 - refSigma0

    # Find the index where the threshold is exceeded beyond 10 m
    mld_idx <- first_valid_index(deltaSigma > densThreshold & P > refDepth)
    if (!is.na(mld_idx)) {
      mld_vec[i] <- P[mld_idx]
    } else {
      mld_vec[i] <- NA
    }
  }
}

# Create a data frame for ggplots and saving plots
df_mld <- data.frame(
  profile = validProfiles,
  value = mld_vec[validProfiles]
)

```

# LIGHT
## Estimation of NA columns
```{r NaN_Light}
col_na_only <- apply(raw_light, 2, function(col) all(is.na(col)))
which(col_na_only)

light_non_nan <- raw_light[, !col_na_only]
# cat("Number of only N/A columns :", sum(col_na_only), "\n\n")
```

## Dark-value correction for light profile
```{r DarkValue}
light_non_log <- exp(raw_light)
light_log <- raw_light
min_window = 4
alpha = 0.01
min_botIntegBound = 150

light_corrected_non_log <- light_non_log

# Find columns that are not entirely empty in `light`
col_na_only <- apply(light_non_log, 2, function(col) all(is.na(col)))
valid_cols <- which(!col_na_only)

delta_for_cstLIGHT <- 1e-8 # define standard delta between min and max of profile to eliminate profiles considered constant
max_depth <- MAX_DEPTH

# Loop over indices of non-empty columns
for (ii_temp in valid_cols) {

  # Initialize local variables
  PCST_temp <- max_depth
  PDARK_temp <- max_depth
  VDARK_temp <- 0

  # Indices of the first and last non-NA values in the profile
  col_temp <- light_non_log[, ii_temp]
  first_non_NaN_temp <- which(!is.na(col_temp))[1]
  last_non_NaN_temp  <- tail(which(!is.na(col_temp)), 1)
  
  cstDetect_temp <- 0

  if ((last_non_NaN_temp - first_non_NaN_temp + 1) < 2) {
    # Not enough valid points to perform a test
    ind_cst_temp <- last_non_NaN_temp + 1
    PCST_temp <- NA
  } else {
    for (jj_temp in first_non_NaN_temp:(last_non_NaN_temp - 1)) {
      vec_temp <- light_corrected_non_log[jj_temp:last_non_NaN_temp, ii_temp]
      # Simple check to avoid processing NAs
      vec_temp <- vec_temp[!is.na(vec_temp)]
      if (length(vec_temp) == 0) {
        range_temp <- NA
      } else {
        # Compute range in the deep part
        range_temp <- diff(range(vec_temp))
      }
      if (is.na(range_temp) || range_temp <= delta_for_cstLIGHT) {
        # Signal considered constant
        cstDetect_temp <- 1
        break # exit loop when range < delta_for_cstLIGHT, i.e., LIGHT signal is constant in interval j:bottom
      }
    }
    # Write index of first value belonging to dark signal
    ind_cst_temp <- jj_temp + 1
    # Write depth of dark value
    PCST_temp <- pres_vec[ind_cst_temp]
  }

  # If constant signal detected, mask the constant part (from bottom upwards)
  if (cstDetect_temp == 1) {
    # Ensure not exceeding max depth
    start_idx <- min(ind_cst_temp + 1, max_depth)
    light_corrected_non_log[start_idx:max_depth, ii_temp] <- NA
    last_non_NaN_temp <- ind_cst_temp - 1
  }

  h_temp <- 1 # By default: signal considered not normal (i.e., not dark)

  if ((last_non_NaN_temp - first_non_NaN_temp + 1) < 4) {
    # Not enough points for a valid test
    ind_dark_temp <- last_non_NaN_temp + 1
    PDARK_temp <- NA
    VDARK_temp <- NA
    slop_temp <- NA
    slopPART_temp <- NA
  } else {
    for (jj_temp in first_non_NaN_temp:(last_non_NaN_temp - 3)) {

      # Sub-sample of the light profile for normality test
      light_segment <- light_corrected_non_log[jj_temp:last_non_NaN_temp, ii_temp]
      light_segment <- light_segment[!is.na(light_segment)]

      # Normality test (Lilliefors ~ lillie.test in package 'nortest')
      if (length(light_segment) >= 4) {
        test_result <- tryCatch(lillie.test(light_segment), error = function(e) NULL)

        # Segment follows normal distribution: considered dark
        if (!is.null(test_result) && test_result$p.value > 0.01) {
          h_temp <- 0  
          break
        }
      }
    }

    # Determine index and depth associated with dark signal
    if (h_temp == 1) {
      # No valid normality test → take the last value
      ind_dark_temp <- last_non_NaN_temp
      PDARK_temp <- pres_vec[ind_dark_temp]
      # VDARK_temp remains 0 (already initialized)
    } else {
      # Normality detected: keep break index
      ind_dark_temp <- jj_temp
      PDARK_temp <- pres_vec[ind_dark_temp]
      VDARK_temp <- median(light_corrected_non_log[ind_dark_temp:last_non_NaN_temp, ii_temp], na.rm = TRUE)
    }

    topdark_temp <- floor(pres_vec[ind_dark_temp])
    botdark_temp <- floor(pres_vec[last_non_NaN_temp]) + 1

    # Compute total attenuation slope
    slop_temp <- (light_log[ind_dark_temp, ii_temp] - light_log[first_non_NaN_temp, ii_temp]) / (pres_vec[ind_dark_temp] - pres_vec[first_non_NaN_temp])

    # Compute partial slope over integration interval (150 = minimum bottom boundary for profiles in linear functional model)
    if (ind_dark_temp > 150) {
      slopPART_temp <- (light_log[150, ii_temp] - light_log[first_non_NaN_temp, ii_temp]) / (pres_vec[min_botIntegBound] - pres_vec[first_non_NaN_temp])
    } else {
      slopPART_temp <- NA
    }
  }

  # LIGHT profile correction

  # Indices to set to NA (unreliable regions)
  toNaN_temp <- unique(c(1:max(1, first_non_NaN_temp - 1), seq(min(ind_dark_temp + 1, MAX_CHLA_DEPTH, last_non_NaN_temp + 1), length(light_corrected_non_log[, ii_temp]))))

  # Remove dark signal if detected (h_temp = 0)
  if (h_temp == 0) {
    light_corrected_non_log[ind_dark_temp:nrow(light_corrected_non_log), ii_temp] <- NA
  }

  # Subtract dark value
  light_corrected_non_log[, ii_temp] <- light_corrected_non_log[, ii_temp] - VDARK_temp

  # Replace negative values with NA
  light_corrected_non_log[light_corrected_non_log[, ii_temp] <= 0, ii_temp] <- NA

  # Linear interpolation if at least 2 valid values
  n_values_temp <- sum(!is.na(light_corrected_non_log[, ii_temp]))
  if (n_values_temp >= 2) {
    vect_temp <- light_corrected_non_log[first_non_NaN_temp:last_non_NaN_temp, ii_temp]
    vect_temp <- zoo::na.approx(vect_temp, na.rm = FALSE)
    light_corrected_non_log[first_non_NaN_temp:last_non_NaN_temp, ii_temp] <- vect_temp
    light_corrected_non_log[toNaN_temp, ii_temp] <- NA
  }
}

# Convert corrected light profile to log scale
light_corrected_log <- log(pmax(light_corrected_non_log, 1e-10))

```

### Quenching depth, based on PAR15
```{r Quenching depth}
# Calculate the depth where the quenching threshold (15 µmol·m⁻²·s⁻¹) is reached (Xing et al., 2018)
quenchDepth_vec <- apply(light_corrected_non_log, 2, function(col) {
  # For each vertical light profile, find the deepest depth where the light value is above 15
  idx <- tail(which(col > 15), 1) # tail = deepest depth
  # If no value exceeds 15, return NA; otherwise, return the index as threshold depth
  if (length(idx) == 0) return(NA) else return(idx)                
})

```

#### Plotting one light profile, before and after dark value correction
```{r LightProfilePlot}
# valid_indices <- !is.na(light_non_log[, idprofile]) | !is.na(light_corrected_non_log[, idprofile])
# xlim_vals <- range(c(light_non_log[, idprofile], light_corrected_non_log[, idprofile]), na.rm = TRUE)
# ylim_vals <- range(pres_vec[valid_indices], na.rm = TRUE)
# 
# plot(light_non_log[, idprofile], pres_vec, type = "l", col = "blue",
#      ylim = rev(ylim_vals), xlim = c(xlim_vals),
#      xlab = "Light ln(µmol/m²/sec)", ylab = "Depth (m)", 
#      main = paste0("Light profile ",idprofile," - Original and Corrected\n", SEAL_NAME, "(",dateTime[idprofile],")"))
# 
# lines(light_corrected_non_log[, idprofile], pres_vec, col = "red")
# legend("bottomright", legend = c("Original", "Corrected"),col = c("blue", "red"), lty = 1, bty = "n")
# 
# abline(h = quenchDepth_vec[idprofile], col = "cyan", lty = 2)
# abline(h = df_mld$value[idprofile], col = "purple", lty = 2)
# 
# legend("bottomleft", legend = c("Quenching Depth", "MLD"), col = c("cyan", "purple"), lty = 1, bty="n")

```

#### Plotting all light profiles, before and after dark value correction
```{r PlotAllProfiles-original&Corrected}
# matplot(raw_light[1:MAX_CHLA_DEPTH,], pres_vec[1:MAX_CHLA_DEPTH],
#         type = "l",
#         ylim = rev(c(0, MAX_CHLA_DEPTH)),
#         lty = 1,
#         xlab = "Light ln(µmol/m²/s)",
#         ylab = "Depth (m)",
#         main = paste0("Light profiles : ", SEAL_NAME))
# grid()
# 
# matplot(light_corrected_log[1:MAX_CHLA_DEPTH,], pres_vec[1:MAX_CHLA_DEPTH],
#         type = "l",
#         xlim = c(-10,10),
#         ylim = rev(c(0, MAX_CHLA_DEPTH)),
#         lty = 1,
#         xlab = "Light ln(µmol/m²/s)",
#         ylab = "Depth (m)",
#         main = paste0("Light profiles corrected : ", SEAL_NAME))
# grid()

```
## Saving light profiles 
```{r SaveAndClassifyLightProfiles, results = 'hide'}
# dir_plot = file.path("oceanographic_data_corrected", DEPLOY_FOLDER, SEAL_NAME, "light")
# dir.create(dir_plot, showWarnings = FALSE, recursive = TRUE)
# 
# nlight = ncol(light_corrected_non_log)
# 
# for (i in 1:nlight) {
#   filename <- paste0(dir_plot, "/", SEAL_NAME, "_", i, ".png")
# 
#   # Ajout d'une exception car probleme de device (dev.off) si les profils à plot sont trop nombreux
#   tryCatch({
#     png(filename, width = 800, height = 600)
# 
#     raw_profile <- light_non_log[, i]
#     corrected_profile <- light_corrected_non_log[, i]
# 
#     # Vérifie s'il y a au moins une valeur non-NA
#     if (all(is.na(raw_profile)) & all(is.na(corrected_profile))) {
#       next  # Skip ce profil si aucune donnée
#     }
#     # Indices valides
#     valid_indices <- !is.na(raw_profile) | !is.na(corrected_profile)
#     # Bornes dynamiques
#     xlim_vals <- range(c(raw_profile, corrected_profile), na.rm = TRUE)
#     ylim_vals <- range(pres_vec[valid_indices], na.rm = TRUE)
# 
#     dateF <- format(dateTime[i], "%d-%m-%Y")
# 
#     plot(raw_profile[1:MAX_CHLA_DEPTH],
#          pres_vec[1:MAX_CHLA_DEPTH],
#          type = "l",
#          col = "blue",
#          ylim = rev(ylim_vals),
#          xlim = xlim_vals,
#          xlab = "Light ln(µmol/m²/sec)",
#          ylab = "Depth (m)",
#          main = paste0("Light profile ", i, " - Original and Corrected (Dark value)\n", SEAL_NAME," (",dateF,")"))
# 
# 
#     lines(corrected_profile[1:MAX_CHLA_DEPTH],
#           pres_vec[1:MAX_CHLA_DEPTH],
#           col = "red")
# 
#     legend("bottomright",
#            legend = c("Original", "Corrected (Dark value)"),
#            col = c("blue", "red"),
#            lty = 1,
#            bty = "n")
# 
#     abline(h = quenchDepth_vec[i], col = "cyan", lty = 2)
#     abline(h = df_mld$value[i], col = "purple", lty = 2)
# 
#     legend("bottomleft",
#            legend = c("Quenching Depth", "MLD"),
#            col = c("cyan", "purple"),
#            lty = 2,
#            bty = "n")
# 
#     dev.off()
#   }, error = function(e) {
#     message("Erreur lors du plot du profil ", i, ": ", conditionMessage(e))
#     if (length(dev.list()) > 0) dev.off()  # ferme si un device est resté ouvert
#   })
# }
# # Ligne pour "évacuer" les devices qui ne sont pas fermés correctement
# while (dev.cur() > 1) dev.off()
```

# FLUO
## Estimation of NA columns
```{r NaN_Chla}
col_na_only <- apply(raw_chla, 2, function(col) all(is.na(col)))
which(col_na_only)

cat("Number of only N/A columns :", sum(col_na_only), "\n\n")
```

## Dark-value correction for Chl-a profiles
```{r DarkValue_chla}
# Input variables
max_depth <- nrow(raw_chla)
PDARK_temp <- MAX_CHLA_DEPTH + 1
delta_for_cstCHLA <- 0.01
darkDepthDelta_temp <- 10

# Extract the last 10 depths from the CHLA matrix
chlaDeep_temp <- raw_chla[(MAX_CHLA_DEPTH - (darkDepthDelta_temp - 1)) : MAX_CHLA_DEPTH, ]

# Median per column (profile)
chlaProfDarkVal_temp <- apply(chlaDeep_temp, 2, median, na.rm = TRUE)

# Fill NAs with the nearest value (note: this line is not identical to Matlab version, watch for differences)
chlaProfDarkVal_temp <- na.approx(chlaProfDarkVal_temp, method = "constant", rule = 2)
chlaProfDarkVal_temp <- round(chlaProfDarkVal_temp, 4) 

# Generate the offset by repeating the median value max_depth times
chlaOffset_temp <- matrix(rep(chlaProfDarkVal_temp, times = max_depth), nrow = max_depth, byrow = TRUE)

darkValProfile <- chlaProfDarkVal_temp

# Subtract dark offset from raw CHLA
CHLA_nadRegDk <- raw_chla
CHLA_nadRegDk <- CHLA_nadRegDk - chlaOffset_temp
CHLA_nadRegDk[CHLA_nadRegDk < 0] <- 0
CHLA_nadRegDk <- round(CHLA_nadRegDk, 4)

# CHLA profiles without columns that are completely NA
chla_non_nan <- CHLA_nadRegDk[, !col_na_only]
n_profile <- ncol(chla_non_nan)

for (ii in 1:n_profile) {
  profile_data <- chla_non_nan[, ii]
  # Profile bounds
  firstNonNan_temp <- which(!is.na(profile_data))[1]
  lastNonNan_temp <- tail(which(!is.na(profile_data)), 1)
  
  if((lastNonNan_temp - firstNonNan_temp + 1) < 2) {
      PDARK_temp <- NaN
  } else {
    for(jj in firstNonNan_temp:(lastNonNan_temp-1)) {
      amplitude_temp <- max(CHLA_nadRegDk[jj:lastNonNan_temp, ii], na.rm = TRUE)
      if(amplitude_temp <= delta_for_cstCHLA) {
        break
      }
    }
    if (jj != lastNonNan_temp - 1) {
      ind_dark_temp <- jj + 1
      PDARK_temp <- pres_vec[ind_dark_temp]
      # If the column contains any values other than NA
      if (!all(is.na(CHLA_nadRegDk[, ii]))) {
        CHLA_nadRegDk[ind_dark_temp:lastNonNan_temp, ii] <- 0
      }
    }
  }
}

# Final corrected CHLA profiles
chla_corrected_dk <- CHLA_nadRegDk

```

### Plotting dark value correction for Chl-a
```{r plot_dk_chla}

# valid_indices <- !is.na(raw_chla[, idprofile]) | !is.na(chla_corrected_dk[, idprofile])
# xlim_vals <- range(c(raw_chla[, idprofile], chla_corrected_dk[, idprofile]), na.rm = TRUE)
# ylim_vals <- range(pres_vec[valid_indices], na.rm = TRUE)
# 
# # Plot only one plot
# plot(raw_chla[1:MAX_CHLA_DEPTH, idprofile], pres_vec[1:MAX_CHLA_DEPTH], type = "l", col = "blue",
#      ylim = rev(ylim_vals),
#      xlim = xlim_vals,
#      xlab = "Chl-a (mg/m³)", ylab = "Depth (m)", main = paste("Chl-a profile (DK)", idprofile, "-", SEAL_NAME, "(",dateTime[idprofile],")"))
# 
# lines(chla_corrected_dk[1:MAX_CHLA_DEPTH, idprofile], pres_vec[1:MAX_CHLA_DEPTH], col = "red")
# 
# legend("bottomright", legend = c("Original", "Corrected"), col = c("blue", "red"), lty = 1, bty = "n")
# 
# # abline(h = quenchDepth_vec[idprofile], col = "cyan", lty = 2)
# abline(h = df_mld$value[idprofile], col = "purple", lty = 2)
# 
# legend("bottomleft", legend = "MLD", col =  "purple", lty = 1, bty="n")
```

### Plotting all dark value correction profiles for Chl-a, before and after correction
```{r plot_all_dk_chla}
# matplot(raw_chla[1:MAX_CHLA_DEPTH,], pres_vec[1:MAX_CHLA_DEPTH],
#         type = "l",
#         ylim = rev(c(0, MAX_CHLA_DEPTH)),
#         lty = 1,
#         xlab = "Chl-a (mg/m³)",
#         ylab = "Depth (m)",
#         main = paste0("Chl-a profiles : ", SEAL_NAME))
# grid()
# 
# matplot(chla_corrected_dk[1:MAX_CHLA_DEPTH,], pres_vec[1:MAX_CHLA_DEPTH],
#         type = "l",
#         ylim = rev(c(0, MAX_CHLA_DEPTH)),
#         lty = 1,
#         xlab = "Chl-a (mg/m³)",
#         ylab = "Depth (m)",
#         main = paste0("Chl-a profiles corrected : ", SEAL_NAME))
# grid()
```

## Non Photochemical Quenching correction for Chl-a profiles (as dark-value correction works only for some profiles, need to verify this code once DK is fixed)
```{r NPQ_correction}
# Xing et al. (2018): 2.1.3 NPQ correction methods based on MLD and light-threshold depth 

CHLA_nadRegDkNpq <- chla_corrected_dk # dk: dark value corrected

# Compute correction depth: min(MLD, quenchDepth)
correction_depth <- ifelse(is.na(mld_vec), quenchDepth_vec,
                         ifelse(is.na(quenchDepth_vec), mld_vec,
                                pmin(mld_vec, quenchDepth_vec)))

# Valid CHLA profiles: at least one non-NA value per column
idx_chlaNonNan <- colSums(!is.na(raw_chla)) > 0
# Valid light profiles
idx_lightNonNan <- colSums(!is.na(raw_light)) > 0

# Loop over valid CHLA and light indices (non-NA)
for (ii in which(idx_chlaNonNan & idx_lightNonNan)) { 

  # Find first index corresponding to min(MLD, QD)
  indCorrection <- which(pres_vec >= correction_depth[ii])[1]

  if (!is.na(indCorrection)) {
    # Find the maximum CHLA value in the interval surface -> correction_depth
    temp_vec <- chla_corrected_dk[1:indCorrection, ii]
    ind_maxChla <- which.max(temp_vec)
    max_chla <- temp_vec[ind_maxChla]
    
    # Correct CHLA: overwrite values from surface to ind_maxChla with max value
    CHLA_nadRegDkNpq[1:ind_maxChla, ii] <- max_chla

    # Update correction depth if an index was found
    if (!is.na(ind_maxChla)) {
      correction_depth[ii] <- ind_maxChla
    }
  }
}

# Final CHLA profiles after NPQ correction
chla_corrected_q <- CHLA_nadRegDkNpq

```

### Flags
```{r}
# solar_angle <- sunAngle(dateTime, lon, lat)$altitude
# n_profile <- ncol(chla_corrected_q)
# flags <- rep("OK", n_profile)
# 
# # NPQ layer might be thinner
# flags[exp(light_corrected_non_log[4,]) < 15] <- "NPQ layer might be thinner than 4m"
# # Shallow mixing conditions
# flags[mld_vec <= quenchDepth_vec] <- "Shallow mixing conditions (MLD shallower than NPQD)"
# # Night profile
# flags[solar_angle <= -12] <- "Night profile"
# # Bottom of npq layer not reached
# flags[is.na(quenchDepth_vec)] <- "No quenching detected"
# # No Light data
# sumNan <- colSums(light_corrected_non_log, na.rm = TRUE)
# flags[sumNan == 0] <- "No Light data"
# # 0 : OK
```

#### Plotting NPQ correction for Chl-a
```{r NPQ_plot}
# valid_indices <- !is.na(chla_corrected_dk[, idprofile]) | !is.na(chla_corrected_q[, idprofile])
# xlim_vals <- range(c(chla_corrected_dk[, idprofile], chla_corrected_q[, idprofile]), na.rm = TRUE)
# ylim_vals <- range(pres_vec[valid_indices], na.rm = TRUE)
# 
# dateF <- format(dateTime[idprofile], "%d-%m-%Y")
# 
# plot(
#   chla_corrected_dk[1:MAX_CHLA_DEPTH, idprofile],
#   pres_vec[1:MAX_CHLA_DEPTH],
#   type = "l", col = "blue",
#   ylim = rev(ylim_vals),
#   xlim = xlim_vals,
#   xlab = "Chl-a  (mg/m³)", ylab = "Depth (m)",
#   main = paste("Profil", idprofile, "-", SEAL_NAME, "(",dateF,") \n flag: ",flags[idprofile])
# )
# 
# lines(chla_corrected_q[1:MAX_CHLA_DEPTH, idprofile], pres_vec[1:MAX_CHLA_DEPTH], col = "red")
# 
# legend("bottomright", legend = c("Original", "Corrected"), col = c("blue", "red"), lty = 1, bty = "n")
# 
# abline(h = quenchDepth_vec[idprofile], col = "cyan", lty = 2)
# abline(h = df_mld$value[idprofile], col = "purple", lty = 2)
# 
# legend("bottomleft", legend = c("Quenching Depth", "MLD"), col = c("cyan", "purple"), lty = 1, bty="n")

```
### Saving Chl-a profiles (Original / Dark value / NPQ)
```{r SaveAndClassifyChlaProfilesFinal, results = 'hide'}
# dir_plot = file.path("oceanographic_data_corrected", DEPLOY_FOLDER, SEAL_NAME,"chla_dk_npq")
# dir.create(dir_plot, showWarnings = FALSE, recursive = TRUE)
# 
# nchla = ncol(raw_chla)
# 
# for (i in 1:nchla) {
#   filename <- paste0(dir_plot, "/", SEAL_NAME, "_DK_NPQ", i, ".png")
#   
#   tryCatch({
#     png(filename, width = 800, height = 600)
#     
#     raw_profile <- raw_chla[, i]
#     corrected_profile <- chla_corrected_q[, i]
#   
#     # Vérifie s'il y a au moins une valeur non-NA
#     if (all(is.na(raw_profile)) & all(is.na(corrected_profile))) {
#       next  # Skip ce profil si aucune donnée
#     }
#     # Indices valides
#     valid_indices <- !is.na(raw_profile) | !is.na(corrected_profile)
#     # Bornes dynamiques
#     xlim_vals <- range(c(raw_profile, corrected_profile), na.rm = TRUE)
#     ylim_vals <- range(pres_vec[valid_indices], na.rm = TRUE)
#     
#     plot(chla_corrected_dk[1:MAX_CHLA_DEPTH, i],
#          pres_vec[1:MAX_CHLA_DEPTH],
#          type = "l",
#          col = "blue",
#          ylim = rev(ylim_vals),
#          xlim = xlim_vals,
#          ylab = "Depth (m)",
#          main = paste0("Chl-a profile ", i, " - Original and Dark Value Corrected\n", SEAL_NAME, " - Correction depth : ", correction_depth[i], " m - flag: ",flags[i], "(",dateTime[i],")"))
#      
#     lines(raw_profile[1:MAX_CHLA_DEPTH],
#           pres_vec[1:MAX_CHLA_DEPTH],
#           col = "red")
#     
#     lines(corrected_profile[1:MAX_CHLA_DEPTH],
#           pres_vec[1:MAX_CHLA_DEPTH],
#           col = "darkgreen")
#     
#     legend("bottomright",
#            legend = c("Original", "Dark value", "NPQ"),
#            col = c("red", "blue", "darkgreen"),
#            lty = 1,
#            bty = "n")
#     
#     abline(h = quenchDepth_vec[i], col = "cyan", lty = 2)
#     abline(h = df_mld$value[i], col = "purple", lty = 2)
#     
#     legend("bottomleft",
#            legend = c("Quenching Depth", "MLD"),
#            col = c("cyan", "purple"),
#            lty = 2,
#            bty = "n")
#     
#     dev.off()
#   }, error = function(e) {
#       message("Erreur lors du plot du profil ", i, ": ", conditionMessage(e))
#       if (length(dev.list()) > 0) dev.off()  # ferme si un device est resté ouvert
#     })
# }
# while (dev.cur() > 1) dev.off()
```

# MAPS
## Bathymetry
```{r Bathy}
min_lon <- 136
max_lon <- 146
min_lat <- -67.3
max_lat <- -65.4

library(raster)

crop <- c(min_lon, max_lon, min_lat, max_lat)
ext_crop <- extent(crop)

large_topo <- raster(BATHY_PATH_LARGE)
large_topo_spat <- terra::rast(large_topo)

large_topo_spat_crop <- crop(large_topo_spat, ext_crop)

large_topo_spat_crop[large_topo_spat_crop >=0] <-NA
```

## Fast ice raster
```{r Fast_ice}
library(raster)
library(terra)
library(dplyr)
library(tidyr)
library(lubridate)
library(zoo)

prj <- "+proj=stere +lat_0=-90 +lat_ts=-70 +datum=WGS84"
ex <- c(-2691055, 2933945, -2390156, 2309844)

r <- rast(FAST_ICE_PATH, "Fast_Ice_Time_Series")
ext(r) <- ext(ex)
crs(r) <- prj
# plot(r)

# define the time needed
# raster_time <- time(r)
# date_min <- as.Date(min(dateTime, na.rm = TRUE))
# date_max <- as.Date(max(dateTime, na.rm = TRUE))
```

```{r}
raster_range <- c(1,2,3,4,5,6,7,8) # Lines to modify to change the SealOfApproval deployment

crop_bounds <- c(130, 150, -67.5, -65)
pts <- cbind(lon = crop_bounds[c(1,2)], lat = crop_bounds[c(3,4)])
prj_pts <- project(pts, to = prj, from = "EPSG:4326")
ex_crop <- ext(prj_pts[2,1], prj_pts[1,1], prj_pts[2,2], prj_pts[1,2])

new_r <- crop(r, ex_crop)
# plot(new_r)

x_sub <- subset(new_r, raster_range)
# plot(x_sub)

#description: Classified surface type: 0 = pack ice or ocean; 1 = continent; 2 = islands; 3 = ice shelf; 4 = fast ice; 5 = manual fast ice edge; 6 = auto fast ice edge.
x_sub_clean <- classify(x_sub, cbind(from = c(0, 2, 5, 6), to = NA))
# persistence
# n <- length(raster_range)
# x_sub_p <- classify(x_sub, cbind(from = c(0, 1, 2, 3, 5, 6), to = NA))
# x_sub_p[x_sub_p == 4] <- 100
# persistence <- sum(x_sub_p, na.rm=TRUE)/n

# projection ...
x_sub_proj <- project(x_sub_clean, crs(large_topo_spat), res = res(large_topo_spat))
# plot(x_sub_proj)

# ... and cropping (dans ce sens là)
x_sub_proj <- crop(x_sub_proj, extent(min_lon, max_lon, min_lat, max_lat))
# plot(x_sub_proj)

# Assign layer names using the actual dates
raster_time <- time(r)[raster_range]
names(x_sub_proj) <- as.character(raster_time)

```

### Fast-ice to dataframe
```{r}
# Convert to long data.frame
x_sub2_df <- as.data.frame(x_sub_proj, xy = TRUE)
x_sub2_long <- pivot_longer(
  x_sub2_df,
  cols = -c(x, y),
  names_to = "date_simple",
  values_to = "value"
) %>%
  mutate(
    month_date = date_simple %>% ymd() %>% + days(5),
    month_label = format(month_date, "%Y - %B"),
    type = factor(month_label, levels = unique(month_label))
  ) %>%
  filter(!is.na(value))  # remove NAs

# Create a new column with descriptive labels
x_sub2_long$surface_type <- factor(x_sub2_long$value,
                                  levels = c(1, 3, 4),
                                  labels = c("continent", "ice_shelf", "fast_ice"))

```

## Mean/sd/median SIC raster map
```{r}
library(raster)
library(terra)
library(stringr)
library(lubridate)
library(dplyr)
library(purrr)
library(tidyr)

# Parameters
wgs <- "+proj=longlat +ellps=WGS84"
amsr_files <- list.files(AMSR_FOLDER_PATH, full.names = TRUE)
# Lines to modify to change the SealOfApproval deployment
target_year_months <- c("202302", "202303", "202304", "202305", "202306", "202307", "202308", "202309") 

crop_bounds <- c(130, 150, -67.5, -65)

# Read a raster to get its CRS
r_test <- raster(amsr_files[1])
crs_native <- projection(r_test)

# Convert WGS84 coordinates to raster native CRS
pts <- cbind(lon = crop_bounds[c(1,2)], lat = crop_bounds[c(3,4)])
pts_sp <- sp::SpatialPoints(pts, proj4string = CRS(wgs))
pts_prj <- sp::spTransform(pts_sp, CRS(crs_native))
prj_coords <- coordinates(pts_prj)

# Define extent for initial crop in native CRS
ex_crop <- extent(min(prj_coords[,1]), max(prj_coords[,1]),
                  min(prj_coords[,2]), max(prj_coords[,2]))

df_SIC <- list()  # Will store data.frames for each month

for (m in target_year_months) {
  selected_files <- amsr_files[str_detect(amsr_files, paste0("s3125-", m))]
  if (length(selected_files) == 0) next

  rasters <- lapply(selected_files, function(f) {
    r <- raster(f)
    r <- setValues(raster(r), values(r))
    crop_r <- crop(r, ex_crop)
    r_proj <- projectRaster(crop_r, crs = wgs)
    final_crop <- extent(min_lon, max_lon, min_lat, max_lat)
    r_cropped <- crop(r_proj, final_crop)
    r_cropped[r_cropped > 100] <- NA_real_
    r_cropped
  })

  r_stack <- stack(rasters)

  mean_r   <- calc(r_stack, mean,   na.rm = TRUE)
  median_r <- calc(r_stack, median, na.rm = TRUE)
  sd_r     <- calc(r_stack, sd,     na.rm = TRUE)

  # Combine all statistics into a single stack
  all_stats <- stack(mean_r, median_r, sd_r)
  names(all_stats) <- c("mean", "median", "sd")

  # Convert to long data.frame
  df_month <- as.data.frame(all_stats, xy = TRUE) %>%
    pivot_longer(
      cols = c(mean, median, sd),
      names_to = "stat",
      values_to = "value"
    ) %>%
    mutate(month_label = format(as.Date(paste0(m, "01"), "%Y%m%d"), "%Y - %B"))

  df_SIC[[m]] <- df_month
}

# Final dataframe
df_SIC <- bind_rows(df_SIC) %>%
  pivot_wider(
    names_from = stat,
    values_from = value
  )
```

# RDS Export
```{r}
# Number of depths and profiles
n_depths <- nrow(raw_light[1:MAX_DEPTH,])
n_profiles <- ncol(raw_light[1:MAX_DEPTH,])

# Transpose matrices to have 1 row per profile
df_light_raw       <- as.data.frame(t(raw_light[1:MAX_DEPTH,])) 
df_light_corrected <- as.data.frame(t(light_corrected_log[1:MAX_DEPTH,]))
df_chla_raw        <- as.data.frame(t(raw_chla[1:MAX_DEPTH,]))
df_chla_corrected  <- as.data.frame(t(chla_corrected_q[1:MAX_DEPTH,]))
df_temperature     <- as.data.frame(t(temperature[1:MAX_DEPTH,]))
df_salinity        <- as.data.frame(t(sal[1:MAX_DEPTH,]))

# Add clear column names (each column = one depth)
depth_names <- paste0("d", seq_len(n_depths))  # d1, d2, ..., dN
colnames(df_light_raw)       <- depth_names
colnames(df_light_corrected) <- depth_names
colnames(df_chla_raw)        <- depth_names
colnames(df_chla_corrected)  <- depth_names
colnames(df_temperature)     <- depth_names
colnames(df_salinity)        <- depth_names

# Add SEAL_NAME column
df_light_raw$seal_id       <- SEAL_NAME
df_light_corrected$seal_id <- SEAL_NAME
df_chla_raw$seal_id        <- SEAL_NAME
df_chla_corrected$seal_id  <- SEAL_NAME
df_temperature$seal_id     <- SEAL_NAME
df_salinity$seal_id        <- SEAL_NAME

# Add profile ID info
df_light_raw$profile_id       <- 1:n_profiles
df_light_corrected$profile_id <- 1:n_profiles
df_chla_raw$profile_id        <- 1:n_profiles
df_chla_corrected$profile_id  <- 1:n_profiles
df_temperature$profile_id     <- 1:n_profiles
df_salinity$profile_id        <- 1:n_profiles

# Convert dateTime to POSIXct for melting
dateTime <- as.POSIXct(dateTime)

# Add date, longitude, latitude
df_light_raw$date <- dateTime
df_light_raw$lon <- lon
df_light_raw$lat <- lat

# Merge all datasets by profile_id
df_all <- merge(df_light_raw, df_light_corrected, by = "profile_id", suffixes = c("_raw", "_corr"))
df_all <- merge(df_all, df_chla_raw, by = "profile_id")
df_all <- merge(df_all, df_chla_corrected, by = "profile_id", suffixes = c("_chla_raw", "_chla_corr"))

# Add date, longitude, latitude again
df_all$date <- df_light_raw$date
df_all$lon <- df_light_raw$lon
df_all$lat <- df_light_raw$lat

# Convert to long format: 1 row = 1 depth × 1 profile
library(reshape2)

# Long format for light
light_raw_long <- melt(df_light_raw,
                       id.vars = c("seal_id", "profile_id", "date", "lon", "lat"),
                       variable.name = "depth_index",
                       value.name = "light_raw")

light_corr_long <- melt(df_light_corrected,
                        id.vars = c("seal_id", "profile_id"),
                        variable.name = "depth_index",
                        value.name = "light_corrected")

chla_raw_long <- melt(df_chla_raw,
                      id.vars = c("seal_id", "profile_id"),
                      variable.name = "depth_index",
                      value.name = "chla_raw")

chla_corr_long <- melt(df_chla_corrected,
                       id.vars = c("seal_id", "profile_id"),
                       variable.name = "depth_index",
                       value.name = "chla_corrected")

temp_long <- melt(df_temperature,
                  id.vars = c("seal_id", "profile_id"),
                  variable.name = "depth_index",
                  value.name = "temperature")

sal_long <- melt(df_salinity,
                  id.vars = c("seal_id", "profile_id"),
                  variable.name = "depth_index",
                  value.name = "salinity")

# Merge all long-format data.frames
df_final <- Reduce(function(x, y) merge(x, y, by = c("seal_id", "profile_id", "depth_index")),
                   list(light_raw_long, light_corr_long, chla_raw_long, chla_corr_long, temp_long, sal_long))

# Add real depth (numeric version needed, otherwise depths sort alphabetically in RDS)
pres_vec <- pres_vec[1:MAX_DEPTH]
df_final$depth <- pres_vec[as.integer(gsub("d", "", df_final$depth_index))]
df_final$depth_num <- as.integer(gsub("d", "", df_final$depth_index))
df_final <- df_final[order(df_final$profile_id, df_final$depth_num), ]
df_final$depth <- pres_vec[df_final$depth_num]

# Reorganize columns
df_final <- df_final[, c("seal_id", "profile_id", "date", "lon", "lat", "depth", 
                         "light_raw", "light_corrected", 
                         "chla_raw", "chla_corrected",
                         "temperature", "salinity")]

# Add MLD and QD (1 value per profile)
df_profile_info <- data.frame(
  profile_id = 1:n_profiles,
  MLD = mld_vec,
  QD = quenchDepth_vec
)

# Merge with df_final (long format)
df_final <- merge(df_final, df_profile_info, by = "profile_id")

df_final$date_simple <- format(df_final$date, "%Y-%m-%d")

cols_to_check <- c("light_raw", "light_corrected", 
                   "chla_raw", "chla_corrected",
                   "temperature", "salinity")
# Remove rows where all cols_to_check are NA
df_final <- df_final[!rowSums(is.na(df_final[, cols_to_check])) == length(cols_to_check), ]

library(FNN)

# Fast-ice
# Extract coordinates
coords_final <- df_final[, c("lon", "lat")]
coords_sub <- x_sub2_long[, c("x", "y")]

# get.knnx finds nearest neighbor in df_sub for each point in df_final
nn <- get.knnx(coords_sub, coords_final, k = 1)

# Retrieve corresponding "fast_ice_idx" value
df_final$fast_ice_idx <- x_sub2_long$value[nn$nn.index[, 1]]

# SIC (Sea Ice Concentration)
# Extract coordinates for df_final (profiles)
coords_profiles <- df_final[, c("lon", "lat")]
coords_SIC <- df_SIC[, c("x", "y")]

# Nearest neighbor spatial (1 nearest)
nn_spatial <- get.knnx(coords_SIC, coords_profiles, k = 1)

# Add SIC columns to df_final
df_final$SIC_mean <- df_SIC$mean[nn_spatial$nn.index[, 1]]
df_final$SIC_median <- df_SIC$median[nn_spatial$nn.index[, 1]]
df_final$SIC_sd <- df_SIC$sd[nn_spatial$nn.index[, 1]]

# Export to .rds (difference with RData: RData stores multiple objects, RDS stores 1 object)
saveRDS(df_final, file = paste0(RSD_PATH, "_data.rds"))

```