---
title: "Maps
author: "Claude Cugerone"
date: "2025-05-22"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, include=FALSE}
library(ncdf4) 
library(raster) 
library(ggplot2)
library(dplyr)
library(lubridate)
library(data.table)
library(oce)
library(terra)
library(tidyterra)
library(cmocean)
library(stringr)
library(zoo)
library(stats)
library(sf)
library(ggnewscale)
library(reshape2)

library(lunar)

```

<h2>Global Variables</h2>
```{r variables}
MAX_CHLA_DEPTH <- 175       # define depth at which to consider dark signal (CHLA "absolute zero")
SURFACE_NOT_NAN <- 4:10

DDU_loc=NULL
DDU_loc$lat = -66.66306	
DDU_loc$lon = 140.00083
DDU_loc=as.data.frame(DDU_loc)
# 66°39’47”S, 140°00’03”E
LON_MIN <- 130
LON_MAX <- 150
LAT_MIN <- -68
LAT_MAX <- -65
```

<h2>Paths and NetCDF extraction</h2>
<h4>Paths</h4>
```{r paths}
ROOT_PROJ <- "/home/cactus/Documents/Oceano/M2/LOCEAN/Weddell_seal_fluo"

# DATA_FOLDER <- "oceanographic_data"
# DEPLOY_FOLDER <- "wd11"
# SEAL_NAME <- paste0(DEPLOY_FOLDER,"-682-17")
# SEAL_FILE <- paste0(SEAL_NAME, "_hr2_prof.nc")
# NC_PATH <- file.path(ROOT_PROJ, DATA_FOLDER, DEPLOY_FOLDER, SEAL_FILE)

DATA_FOLDER <- "oceanographic_data_corrected"
DEPLOY_FOLDER <- "wd11"
SEAL_NAME <- paste0(DEPLOY_FOLDER,"-913-18")
SEAL_FILE <- paste0(SEAL_NAME, "_data.rds")
RDS_PATH <- file.path(ROOT_PROJ, DATA_FOLDER, DEPLOY_FOLDER, SEAL_FILE)

DEPLOYMENT_FILE <- paste0("combined_",DEPLOY_FOLDER,".rds")
RDS_DEPLOYMENT_PATH <- file.path(ROOT_PROJ, DATA_FOLDER, DEPLOY_FOLDER, DEPLOYMENT_FILE)

BATHY_FOLDER <- "bathy"
FINE_TOPO_FILE <- "gvdem100v3/w001001.adf"
LARGE_TOPO_FILE <- "gebco_2023_n-50.0_s-79.0_w90.0_e180.0.tif"
BATHY_PATH_FINE <- file.path(ROOT_PROJ, BATHY_FOLDER, FINE_TOPO_FILE)
BATHY_PATH_LARGE <- file.path(ROOT_PROJ, BATHY_FOLDER, LARGE_TOPO_FILE)

FAST_ICE_FOLDER <- "fast_ice"
FAST_ICE_FILE <- "mertz_sara_akiko_19.nc"
FAST_ICE_PATH <- file.path(ROOT_PROJ, FAST_ICE_FOLDER, FAST_ICE_FILE)

AMSR_FOLDER <- "amsr2"
AMSR_FOLDER_YEAR <- "2019"
AMSR_FOLDER_PATH <- file.path(ROOT_PROJ,AMSR_FOLDER,AMSR_FOLDER_YEAR)
AMSR_FILE <- "asi-AMSR2-s3125-20190401-v5.4.tif"
AMSR_PATH <- file.path(ROOT_PROJ,AMSR_FOLDER,AMSR_FOLDER_YEAR,AMSR_FILE)
```

<h2>MAPS</h2>
<h4>Reading RDS</h4>
```{r}
# RDS qui contient les valeurs de lumieres et chla corrigées par le script Data_L_F_corrected
df_data_corrected<- readRDS(RDS_PATH)

# df_data_corrected contient une ligne par profondeur et par profil
# Il faut reformatter : profondeur = lignes, profil = colonnes
light_corrected_log <- dcast(df_data_corrected, depth ~ profile_id, value.var = "light_corrected")

# light_corrected_log contient la colonne "depth", on peut l'enlever
pres_vec <- light_corrected_log$depth
light_corrected_log <- as.matrix(light_corrected_log[ , -which(colnames(light_corrected_log) == "depth")])

df_profiles_mld <- df_data_corrected[!duplicated(df_data_corrected$profile_id), c("profile_id", "MLD")]
mld_vec <- df_profiles_mld$MLD

df_profiles_qd <- df_data_corrected[!duplicated(df_data_corrected$profile_id), c("profile_id", "QD")]
quenchDepth_vec <- df_profiles_qd$QD

df_profiles_lon <- df_data_corrected[!duplicated(df_data_corrected$profile_id), c("profile_id", "lon")]
lon <- df_profiles_lon$lon

df_profiles_lat <- df_data_corrected[!duplicated(df_data_corrected$profile_id), c("profile_id", "lat")]
lat <- df_profiles_lat$lat

df_profiles_date <- df_data_corrected[!duplicated(df_data_corrected$profile_id), c("profile_id", "date")]
date <- df_profiles_date$date
```

<h4>Loading bathymetry</h4>
```{r Bathy}
fine_topo <- raster(BATHY_PATH_FINE)
fine_topo[fine_topo >=0] <- NA
fine_topo_spat <- terra::rast(fine_topo)

large_topo <- raster(BATHY_PATH_LARGE)
large_topo[large_topo >=0] <-NA
large_topo_spat <- terra::rast(large_topo)

fine_topo_spat <- crop(fine_topo_spat, extent(LON_MIN, LON_MAX, LAT_MIN, LAT_MAX))
large_topo_spat <- crop(large_topo_spat, extent(LON_MIN, LON_MAX, LAT_MIN, LAT_MAX))
```

<h4>Loading and cropping fast ice raster</h4>
```{r Fast_ice}
prj <- "+proj=stere +lat_0=-90 +lat_ts=-70 +datum=WGS84"
ex <- c(-2691055, 2933945, -2390156, 2309844)

fast_ice_raster <- rast(FAST_ICE_PATH, "Fast_Ice_Time_Series")
set.ext(fast_ice_raster, ext(ex))
set.crs(fast_ice_raster, crs(prj))
#plot(fast_ice_raster[[1]], legend = FALSE) # le continent

points <- cbind(lon = c(LON_MIN, LON_MAX), lat = c(LAT_MIN, LAT_MAX))

prjpoints <- project(points, to = prj, from = "EPSG:4326")
new_ex<-extent(prjpoints[2,1],prjpoints[1,1],prjpoints[2,2],prjpoints[1,2])
new_r=crop(fast_ice_raster,new_ex)
# From feb to may
time_index <- seq(1,4, by = 1) 
x_sub <- subset(new_r, time_index)
plot(x_sub, legend = FALSE)
```

<h4>Removing values from the raster we don't want </h4>
```{r fast_ice_processing}
x_sub2= project(x_sub, crs(large_topo_spat), 
                res = res(large_topo_spat)) 

# Create a sequence of months between the first and last month
start <- format(as.Date("2019-02-05"), "%m")
end <- format(as.Date("2019-05-06"), "%m")

months_seq <- seq(from = start, to = end)

# Extract the month names from the sequence
month_names <- paste0("2019 - ", month.name[months_seq])

names(x_sub2) <- month_names
plot(x_sub2, legend = FALSE)
x_sub2 <- crop(x_sub2, extent(LON_MIN, LON_MAX, LAT_MIN, LAT_MAX))
plot(x_sub2, legend = FALSE)

#description: Classified surface type: 0 = pack ice or ocean; 1 = continent; 2 = islands; 3 = ice shelf; 4 = fast ice; 5 = manual fast ice edge; 6 = auto fast ice edge.
x_sub2=round(x_sub2)

# continent <- x_sub2
# continent[x_sub2 != 1] <- NA
# continent[x_sub2 == 1] <- 1
# 
# glacier <- x_sub2
# glacier[x_sub2 != 4] <- NA
# glacier[x_sub2 == 4] <- 1
# 
# ice_shelf <- x_sub2
# ice_shelf[x_sub2 != 3] <- NA
# ice_shelf[x_sub2 == 3] <- 1

x_sub2_df <- as.data.frame(x_sub2, xy = TRUE)
x_sub2_long <- pivot_longer(x_sub2_df, cols = starts_with("2019"), names_to = "month_label", values_to = "value")
x_sub2_long_select <- x_sub2_long # %>% filter(value == c(1,3,4))

# x_sub2_wide <- x_sub2_long %>%
#   pivot_wider(names_from = month, values_from = value)
# 
# x_sub2_wide_filtered <- x_sub2_wide %>%
#   filter(if_any(starts_with("2019"), ~ .x %in% c(1,3,4)))

```

<h4>Loading and cropping AMSR raster</h4>
```{r AMSR}
# wgs <- "+proj=longlat +ellps=WGS84"
#  amsr_raster <- raster(AMSR_PATH)
#  amsr_raster[amsr_raster > 100] <- NA_real_
#  amsr_raster <- setValues(raster(amsr_raster), values(amsr_raster))
#  plot(amsr_raster)
#  
#  points <- cbind(lon = c(LON_MIN, LON_MAX), lat = c(LAT_MIN, LAT_MAX))
#  
#  boxext <- extent(projectExtent(raster(extent(c(LON_MIN,LON_MAX,LAT_MIN,LAT_MAX)), crs = wgs), projection(amsr_raster)))
#  amsr_cropped=crop(amsr_raster, boxext)
#  plot(amsr_cropped)
#  
#  amsr_raster_cropped <- projectRaster(amsr_cropped, crs = wgs)
#  plot(amsr_raster_cropped)
```

<h4>Computing mean/sd/median of SIC raster map, and plotting</h4>
```{r AMSR_processing}

wgs <- "+proj=longlat +ellps=WGS84"
all_files <- list.files(AMSR_FOLDER_PATH, full.names = TRUE)

# on cible une date et un mois, pour ensuite en calculer la moyenne, mediane et sd
target_year_month <- "201903"
date_obj <- as.Date(paste0(target_year_month, "01"), format = "%Y%m%d")
formatted_title <- format(date_obj, "%B %Y") 

# on selectionne tous les fichiers du mois
selected_files <- all_files[str_detect(all_files, paste0("s3125-", target_year_month))]

raster_list <- lapply(selected_files, function(f) {
  r <- raster(f)
  r[r > 100] <- NA_real_
  r
})

raster_stack <- stack(raster_list)
ref_raster <- raster_list[[1]]

points <- cbind(lon = c(LON_MIN, LON_MAX), lat = c(LAT_MIN, LAT_MAX))

# Recadrage
boxext <- extent(projectExtent(raster(extent(c(LON_MIN, LON_MAX, LAT_MIN, LAT_MAX)), crs = wgs), projection(ref_raster)))
raster_cropped <- crop(raster_stack, boxext)

mean_raster <- calc(raster_cropped, fun = mean, na.rm = TRUE)
median_raster <- calc(raster_cropped, fun = median, na.rm = TRUE)
sd_raster <- calc(raster_cropped, fun = sd, na.rm = TRUE)

mean_raster_wgs <- projectRaster(mean_raster, crs = wgs)
median_raster_wgs <- projectRaster(median_raster, crs = wgs)
sd_raster_wgs <- projectRaster(sd_raster, crs = wgs)

mean_spat <- rast(mean_raster_wgs)
median_spat <- rast(median_raster_wgs)
sd_spat <- rast(sd_raster_wgs)

plot(mean_spat, main = paste("AMSR mean -", formatted_title))
plot(median_spat, main = paste("AMSR mediane -", formatted_title))
plot(sd_spat, main = paste("AMSR  SD -", formatted_title))
```
<h4>Fusion of SIC fast ice and light profiles</h4>
```{r Draft map SIC FI}
# Trouver une meilleure façon d'automatiser quelle mois on affiche, si c'est avec mean sd ou median, light ou chla

# month_profile = 3 #
# profiles_per_month <- df_light %>%
#   filter(lubridate::month(date) == month_profile) %>%
#   mutate(
#     month = lubridate::month(date),
#     mld = mld_vec[1:nrow(.)],  # Assurer que mld_vec a la même longueur que le subset
#     qd = quenchDepth_vec[1:nrow(.)]  # Pareil pour quenchDepth_vec
#   )
# 
# continent_factor <- as.factor(continent[[1]])  # Convertit la couche en facteur
# glacier_factor <- as.factor(glacier[[1]])  # Convertit la couche en facteur
# 
# ggplot() +
#   geom_spatraster_contour(data = fine_topo_spat) + # bathy
#   geom_spatraster_contour_text(data = fine_topo_spat) +
#   
#   geom_spatraster(data = mean_spat, alpha = 0.8) + # moyenne AMSR2 sur le mois
#   scale_fill_viridis_c(na.value = NA) +  # NA rendus transparents
#   
#   new_scale_fill() +
#   geom_spatraster(data = glacier_factor, show.legend = FALSE) +
#   scale_fill_manual(values = c("1" = "#A9D6E5"), na.value = NA) +  # Continent en gris, NA en blanc
#   
#   new_scale_fill() +
#   geom_spatraster(data = continent_factor, show.legend = FALSE) +
#   scale_fill_manual(values = c("1" = "slategray"), na.value = NA) +  # Continent en gris, NA en blanc
#   
#   geom_point(data = profiles_per_month, aes(x = lon, y = lat), shape = 21, color = "black", fill = "yellow", size = 1) +
#   
#   geom_point(aes(x = DDU_loc$lon, y = DDU_loc$lat), color = "red", shape = 4, size = 3, stroke = 1) +  # DDU
#   annotate("text", x = DDU_loc$lon, y = DDU_loc$lat, label = "DDU", hjust = -0.2, vjust = -0.5, color = "red", size = 4) +  # DDU label
#   
#   xlab("Longitude") + ylab("Latitude") +
#   theme_minimal() 
```

<h4>Fusion of SIC fast ice and light profiles (smaller crop)</h4>
```{r SIC_fast_icePlot}

# Même remarque qu'au dessus

# continent_factor <- as.factor(continent[[1]])  # Convertit la couche en facteur
# glacier_factor <- as.factor(glacier[[1]])
# ice_shelf_factor <- as.factor(ice_shelf[[1]])
# 
# ggplot() +
#   geom_spatraster_contour(data = fine_topo_spat, color = "black") + # bathy
#   geom_spatraster_contour_text(data = fine_topo_spat)+
#   
#   new_scale_fill() +
#   geom_spatraster(data = mean_spat, alpha = 0.8) + # moyenne AMSR2 sur le mois
#   scale_fill_viridis_c(na.value = NA) +  # NA rendus transparents
#   
#   new_scale_fill() +
#   geom_spatraster(data = ice_shelf_factor, show.legend = FALSE) +
#   scale_fill_manual(values = c("1" = "#A9D6E5"), na.value = NA) +  # ice shelf, NA en blanc
#   
#   
#   new_scale_fill() +
#   geom_spatraster(data = glacier_factor,alpha = 0.9, show.legend = FALSE) +
#   scale_fill_manual(values = c("1" = "#A9D6E5"), na.value = NA) +  # glacier, NA en blanc
#   
#   new_scale_fill() +
#   geom_spatraster(data = continent_factor,alpha = 0.4, show.legend = FALSE) +
#   scale_fill_manual(values = c("1" = "slategray"), na.value = NA) +  # Continent en gris, NA en transparent
#   
#   new_scale_fill() +
#   geom_point(data = profiles_per_month, aes(x = lon, y = lat, color = light_surf), size = 1) +
#   scale_color_viridis_c(limits = c(0.05, 0.1), oob = scales::squish, na.value = NA) +
#   
#   geom_point(aes(x = DDU_loc$lon, y = DDU_loc$lat), color = "red", shape = 4, size = 3, stroke = 1) +  # DDU
#   annotate("text", x = DDU_loc$lon, y = DDU_loc$lat, label = "DDU", hjust = -0.2, vjust = -0.5, color = "red", size = 4) +  # DDU label
#   
#   xlab("Lon") + ylab("Lat") +
#   ggtitle(paste0("Bathy, ice and light values close to DDU (", profiles_per_month$month, ")")) +
#   theme_minimal() +
#   coord_sf(xlim = c(139.5, 145),ylim = c(-67, -65.5))

```

```{r}

# C'est la taille de la carte que l'on souhaite 
ext_crop <- ext(139.5, 140.5, -66.8, -66.4)

# Crop raster
mean_spat_crop <- crop(mean_spat, ext_crop)
fine_topo_spat_crop <- crop(fine_topo_spat, ext_crop)

# Crop dataset fast-ice ice shelf continent 
x_sub2_long_crop <- x_sub2_long_select %>%
  dplyr::filter(x >= 139.5, x <= 140.5,
                y >= -66.8, y <= -66.4)

```

<h4>Carte LIGHT avec tous les individus </h4>
```{r}
df_deployment_data <- readRDS(RDS_DEPLOYMENT_PATH)

light_at_4m <- subset(df_deployment_data, depth == 4)
light_at_4m <- light_at_4m[!is.na(light_at_4m$light_corrected), ]
light_at_4m$month <- format(light_at_4m$date, "%m")

plot_light <- light_at_4m[, c("date", "lon", "lat", "light_corrected", "month")]
plot_light <- plot_light[order(plot_light$month), ]

plot_light$month_label <- format(plot_light$date, "%Y - %B")
plot_light$month_label <- factor(plot_light$month_label, 
    levels = unique(format(seq(min(plot_light$date), max(plot_light$date), by = "month"), "%Y - %B")))

# Tester la carte
ggplot() +
  geom_spatraster_contour(data = fine_topo_spat_crop, color = "black") +
  #geom_spatraster_contour_text(data = fine_topo_spat_crop) +

  new_scale_fill() +
  geom_spatraster(data = mean_spat_crop, alpha = 0.7) +
  scale_fill_cmocean(name = "ice",na.value = NA) +

  new_scale_fill() +
  geom_raster(data = x_sub2_long_crop, aes(x = x, y = y, fill = factor(value)), alpha = 0.8, show.legend = FALSE) +
  scale_fill_manual(values = c(
  "1" = "slategray",  # continent
  "3" = "#A9D6E5",    # ice shelf
  "4" = "#A9D6E5"     # glacier
  ), na.value = NA) +

  new_scale_color() +
  geom_point(data = plot_light, aes(x = lon, y = lat, color = light_corrected), size = 1) +
  scale_color_viridis_c(limits = c(-15, 5), oob = scales::squish, na.value = NA) +
  
  geom_point(data = DDU_loc, aes(x = lon, y = lat), color = "red", shape = 4, size = 2, stroke = 1)+
  annotate("text", x = DDU_loc$lon, y = DDU_loc$lat, label = "DDU", hjust = -0.2, vjust = -0.5, color = "red", size = 4) +
  
  coord_sf(xlim = c(139.5, 140.5), ylim = c(-66.8, -66.4)) +
  facet_wrap(~month_label) +
  theme_minimal() +
  labs(x = "Lon", y = "Lat", title = paste0(DEPLOY_FOLDER, "- Light values at 4 m deep by month"))

```

<h4>Carte CHLA avec tous les individus </h4>
```{r}
df_deployment_data <- readRDS(RDS_DEPLOYMENT_PATH)

chla_at_4m <- subset(df_deployment_data, depth == 4)
chla_at_4m <- chla_at_4m[!is.na(chla_at_4m$chla_corrected), ]
chla_at_4m$month <- format(chla_at_4m$date, "%m")

plot_chla <- chla_at_4m[, c("date", "lon", "lat", "chla_corrected", "month")]
plot_chla <- plot_chla[order(plot_chla$month), ]

plot_chla$month_label <- format(plot_chla$date, "%Y - %B")
plot_chla$month_label <- factor(plot_chla$month_label, 
    levels = unique(format(seq(min(plot_chla$date), max(plot_chla$date), by = "month"), "%Y - %B")))

# Tester la carte
ggplot() +
  geom_spatraster_contour(data = fine_topo_spat_crop, color = "black") +
  #geom_spatraster_contour_text(data = fine_topo_spat_crop) +

  new_scale_fill() +
  geom_spatraster(data = mean_spat_crop, alpha = 0.7) +
  scale_fill_cmocean(name = "ice",na.value = NA) +

  new_scale_fill() +
  geom_raster(data = x_sub2_long_crop, aes(x = x, y = y, fill = factor(value)), alpha = 0.8, show.legend = FALSE) +
  scale_fill_manual(values = c(
  "1" = "slategray",  # continent
  "3" = "#A9D6E5",    # ice shelf
  "4" = "red"     # fast-ice
  ), na.value = NA) +

  new_scale_color() +
  geom_point(data = plot_chla, aes(x = lon, y = lat, color = chla_corrected), size = 1) +
  scale_color_viridis_c(limits = c(0, 15), oob = scales::squish, na.value = NA) +
  
  geom_point(data = DDU_loc, aes(x = lon, y = lat), color = "red", shape = 4, size = 2, stroke = 1)+
  annotate("text", x = DDU_loc$lon, y = DDU_loc$lat, label = "DDU", hjust = -0.2, vjust = -0.5, color = "red", size = 4) +
  
  coord_sf(xlim = c(139.5, 140.5), ylim = c(-66.8, -66.4)) +
  facet_wrap(~month_label) +
  #  facet_wrap(~month_label(c("", "", "")))
  theme_minimal() +
  labs(x = "Lon", y = "Lat", title = paste0(DEPLOY_FOLDER, "- Chlorophyl values at 4 m deep by month"))

```
<h4>Dataframe creation for analyzing profiles position</h4>
```{r}
depth_interval <- pres_vec
light_interval <- light_corrected_log

# fonction moyenne sur un intervalle de profondeur (non pondérée)
simple_light_avg <- function(depths, lights, min_depth, max_depth) {
  idx <- which(depths >= min_depth & depths <= max_depth)
  mean(lights[idx], na.rm = TRUE)
}

# Moyenne simple sur la tranche 4–10 m
light_surf_avg <- apply(light_interval, 2, function(x) {
  simple_light_avg(depth_interval, x, min_depth = 4, max_depth = 10)
})

# Calcul angle solaire et bins
solar_angle <- sunAngle(date, lon, lat)$altitude
solar_threshold <- -12
lat_bin <- floor(lat * 20) / 20

df_light <- data.frame(
  date = date,
  lon = lon,
  lat = lat,
  lat_bin = lat_bin,
  lat_bin_label = factor(paste0("[", lat_bin, "°, ", lat_bin + 0.05, "°[")),
  solar_angle = solar_angle,
  light_surf = light_surf_avg
)

# Ajouter la période jour/crépuscule/nuit
df_solar <- df_light %>%
  mutate(
    period = case_when(
      solar_angle >= 0 ~ "Day",
      solar_angle < 0 & solar_angle > solar_threshold ~ "Twilight",
      solar_angle <= solar_threshold ~ "Night"
    ))

nPeriod = df_solar %>%  count(period)

cat("Number of Day values :", nPeriod$n[1], "\n")
cat("Number of Night values :", nPeriod$n[2], "\n")
cat("Number of Twilight values :", nPeriod$n[3], "\n")

# Ajouter la phase lunaire
df_solar$lunar_phase <- lunar.phase(df_solar$date, shift=12, name = TRUE)

# Extraire les dates de pleine lune (distinctes, arrondies à la date)
full_moon_dates <- df_solar %>%
  filter(lunar_phase == "Full") %>%
  mutate(date_only = as.Date(date)) %>%
  distinct(date_only) %>%
  pull(date_only) %>%
  as.POSIXct(tz = "UTC")

```

<h4>Computing solar angle & classifying in Day / Night / Twilight by atitude bins</h4>
```{r SolarAngle}

# Angle solaire par tranche de latitude
p1 <- ggplot(df_solar, aes(x = date, y = solar_angle, color = period)) +
  geom_point() +
  facet_wrap(~lat_bin_label, scales = "fixed") +
  coord_cartesian(ylim = c(-40, 40)) +
  labs(
    title = "Solar Angle by Latitude Bin",
    x = "Date", y = "Solar Angle (°)"
  ) +
  scale_color_manual(values = c("Day" = "red", "Twilight" = "green", "Night" = "navy")) +
  theme_minimal()

# Lumière par tranche de latitude
p2 <- ggplot(df_solar, aes(x = date, y = light_surf, color = period)) +
  geom_point() +
  geom_vline(xintercept = full_moon_dates, linetype = "dashed", color = "black") +
  facet_wrap(~lat_bin_label, scales = "free_y") +
  coord_cartesian() +
  labs(
    title = paste0("Light (", SURFACE_NOT_NAN[1], " to ", SURFACE_NOT_NAN[7], "m deep) by Latitude Bin"),
    x = "Date", y = "ln(µmol/m²/sec)"
  ) +
  scale_color_manual(values = c("Day" = "red", "Twilight" = "green", "Night" = "navy")) +
  theme_minimal()

print(p1)
print(p2)

```

<h4>Plotting solar angle and light values at 4-10m deep</h4>
```{r LightThroughTime}
# Plot angle solaire
ggplot(df_solar, aes(x = date, y = solar_angle, color = period)) +
  geom_point() + labs(title = "Solar Angle", x = "Date", y = "Solar Angle (°)") +
  scale_color_manual(values = c("Day" = "red", "Twilight" = "green", "Night" = "navy")) +
  theme_minimal()

# Plot lumière avec annotations des pleines lunes
ggplot(df_solar, aes(x = date, y = light_surf, color = period)) +
  geom_point() +
  geom_vline(xintercept = full_moon_dates, linetype = "dashed", color = "black") +
  labs(title = paste0("Light (", SURFACE_NOT_NAN[1], " to ", SURFACE_NOT_NAN[7], "m deep)"), x = "", y = "ln(µmol/m²/sec)") +
  scale_color_manual(values = c("Day" = "red", "Twilight" = "green", "Night" = "navy")) +
  theme_minimal()


# Waxing : croissant - Waning : décroissante
ggplot(df_solar, aes(x = date, y = solar_angle, color = lunar_phase)) +
  geom_point() +
  geom_vline(xintercept = full_moon_dates, linetype = "dashed", color = "black") +
  labs(
    title = paste0("Light (", SURFACE_NOT_NAN[1], " to ", SURFACE_NOT_NAN[7], "m) by Lunar Phase"),
    x = "", y = "ln(µmol/m²/sec)"
  ) +
  scale_color_brewer(palette = "Dark2") +
  theme_minimal()

ggplot(df_solar, aes(x = date, y = light_surf, color = lunar_phase)) +
  geom_point() +
  geom_vline(xintercept = full_moon_dates, linetype = "dashed", color = "black") +
  labs(
    title = paste0("Light (", SURFACE_NOT_NAN[1], " to ", SURFACE_NOT_NAN[7], "m) by Lunar Phase"),
    x = "", y = "ln(µmol/m²/sec)"
  ) +
  scale_color_brewer(palette = "Dark2") +
  theme_minimal()

```

<h4>Geographical Classification of Profiles Relative to the DDU Station</h4>
```{r Light_profile_NSEW}
df_solar$lon <- lon
df_solar$lat <- lat

lon_center <-  DDU_loc$lon 
lat_center <-  DDU_loc$lat

df_solar <- df_solar %>%
  mutate(zone = case_when(
    lon <= lon_center & lat >= lat_center ~ "Nord-Ouest",
    lon > lon_center & lat >= lat_center ~ "Nord-Est",
    lon <= lon_center & lat < lat_center ~ "Sud-Ouest",
    lon > lon_center & lat < lat_center ~ "Sud-Est"
  ))

# A revoir (on a supprimé ice_shelf_factor)
ggplot() +
  # bathy et continent
  geom_spatraster_contour(data = fine_topo_spat) +  # lignes de contour
  geom_spatraster_contour_text(data = fine_topo_spat) +  # texte des courbes
  
  geom_spatraster(data = ice_shelf_factor, show.legend = FALSE) +
  scale_fill_manual(values = c("1" = "#A9D6E5"), na.value = NA) +  # ice shelf
  
  new_scale_fill() +
  geom_spatraster(data = continent_factor, show.legend = FALSE) +
  scale_fill_manual(values = c("1" = "slategray"), na.value = NA) +  # continent
  
  new_scale_fill() +
  geom_spatraster(data = glacier_factor,alpha = 0.9, show.legend = FALSE) +
  scale_fill_manual(values = c("1" = "#A9D6E5"), na.value = NA) +  # glacier
  
  geom_point(data = df_solar, aes(x = lon, y = lat, color = zone, shape = period), size = 3, alpha = 0.8) +   # Données solaires

  geom_point(aes(x = DDU_loc$lon, y = DDU_loc$lat), color = "red", shape = 4, size = 3, stroke = 1) +
  annotate("text", x = DDU_loc$lon, y = DDU_loc$lat, label = "DDU", hjust = -0.2, vjust = -0.5, color = "red", size = 4) +  # DDU

  labs(title = "Zones (couleur) et période jour/nuit/crépuscule (forme)", x = "Longitude", y = "Latitude") +
  scale_color_manual(values = c("Nord-Ouest" = "darkorange", "Nord-Est" = "purple", "Sud-Ouest" = "blue", "Sud-Est" = "darkgreen")) +
  scale_shape_manual(values = c("Day" = 16, "Twilight" = 17, "Night" = 15)) +
  theme_minimal() +
  coord_sf(xlim = c(139, 140.5),ylim = c(-66.8, -66.4))
```

<h4>Classification of Profiles Based on Their Distance from the DDU Station</h4>
```{r LightProfile_distance_DDU}
library(geosphere)

df_solar <- df_solar %>%
  mutate(distance_to_center_km = distHaversine(cbind(lon, lat), c(lon_center, lat_center)) / 1000)  # en km

df_solar <- df_solar %>%
  mutate(distance_group = case_when(
    distance_to_center_km < 10 ~ "Centre",
    distance_to_center_km < 25 ~ "Périphérie proche",
    TRUE ~ "Périphérie lointaine"
  ))

ggplot() +
  # bathy et continent
  geom_spatraster_contour(data = fine_topo_spat) +
  geom_spatraster_contour_text(data = fine_topo_spat) +
  
  geom_spatraster(data = ice_shelf_factor, show.legend = FALSE) +
  scale_fill_manual(values = c("1" = "#A9D6E5"), na.value = NA) +  # ice shelf
  
  new_scale_fill() +
  geom_spatraster(data = continent_factor, show.legend = FALSE) +
  scale_fill_manual(values = c("1" = "slategray"), na.value = NA) +  # continent
  
  new_scale_fill() +
  geom_spatraster(data = glacier_factor,alpha = 0.9, show.legend = FALSE) +  # glacier
  scale_fill_manual(values = c("1" = "#A9D6E5"), na.value = NA) +

  geom_point(data = df_solar, aes(x = lon, y = lat, color = distance_group, shape = period), size = 2, alpha = 0.8) +  # Groupes de distance

  geom_point(aes(x = DDU_loc$lon, y = DDU_loc$lat), color = "red", shape = 4, size = 3, stroke = 1) +
  annotate("text", x = DDU_loc$lon, y = DDU_loc$lat, label = "DDU", hjust = -0.2, vjust = -0.5, color = "red", size = 4) +  # DDU

  labs(title = "", x = "Longitude", y = "Latitude") + scale_color_manual(values = c("Centre" = "firebrick", "Périphérie proche" = "goldenrod", "Périphérie lointaine" = "steelblue")) +
  scale_shape_manual(values = c("Day" = 16, "Twilight" = 17, "Night" = 15)) +
  theme_minimal() +
  coord_sf(xlim = c(139, 140.5),ylim = c(-66.8, -66.4))
```
